---
title: "Hypothesis Tests"
author: "Vladislav Kargin"
date: "October 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="50%")
```


For these illustrations, we will use the data from the book by Chihara and Hestenberg "Mathematical Statistics with Resampling and R". The data available at https://sites.google.com/site/chiharahesterberg/data2 or https://www2.math.binghamton.edu/p/people/kargin/math448/start


## One-sample t-test

(The original data was from https://www.bgs.ac.uk/geological-data/datasets/. However, it is not available there anymore)

Here is the example:

A study recommends that the maximum concentration of arsenic in irrigation water be 100 ðœ‡g/l to prevent a buildup of the chemical that might harm future crop production. How does the water in Bangladesh measure up to this recommendation? Let ðœ‡ denote the true mean level of arsenic in Bangladesh wells. We wish to test H0 âˆ¶ ðœ‡ = 100 versus HA âˆ¶ ðœ‡ > 100.

```{r}
Bangladesh <- read.csv("./data/bangladesh.csv")

head(Bangladesh)
summary(Bangladesh)


result <- t.test(Bangladesh$Arsenic, mu = 100, alt = "greater", conf.level = 0.95)
result
```



Note that p-value is 0.08, so we cannot reject the null hypothesis at $5\%$ level. We can also see this conclusion from the fact that the (one-sided) confidence interval contains the value $100$. 

## Two-sample t-test


The Centers for Disease Control and Prevention (CDC) maintains a database on all babies born in a given year (http://wonder.cdc.gov/natality-current.html).One data set that we will investigate consists of a random sample of 1009 babies born in North Carolina during 2004.

Is there a real difference in the mean weights of North Carolina babies born to nonsmoking and smoking mothers in 2004?

```{r}
data <- read.csv("./data/NCbirths2004.csv")
head(data)
summary(data)

babies <- as.data.frame(data, stringsAsFactors = TRUE)  # Convert all columns to factor
              
head(babies)
summary(babies)

t.test(Weight ~ Tobacco, data = babies, alt = "greater", var.equal = F) #we do not assume that variances are the same, which is the default behavior. 
```

p-value is very small, so we can reject the null hypothesis at $\alpha = 0.01$ and conclude that there is a real difference in baby weights for babies born to nonsmoking and smoking mothers. 

Inciedentally, we can test whether the variances are the same. 

```{r}
var.test(Weight ~ Tobacco, data = babies) 
```

p-value indicates that we cannot reject the hypothesis that variances are the same. 

### Two-sample test for proportions

Do men and women differ in their beliefs about an afterlife? In the 2002 General Social Survey (see case study in Section 1.7 of Chihara-Hestenberg book), participants were asked this question, and of the 684 women who responded, 550 said yes (80.40%), compared to 425 of the 563 men (75.49%).

```{r}
prop.test(c(550, 425), c(684, 563), correct = TRUE) #this is done with continuity correction (default behavior)

```
According to the p-value, we can reject the null hypothesis at $5\%$ significance level but not at $1\%$ level. Some supportive evidence exists but it is not very strong. 




