---
title: "Math448 HW4"
author: "your name"
date: "December 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Infection Risk in Hospitals 


Data from $n = 113$ hospitals in the United States were collected in order to assess factors related to the likelihood that a hospital patient acquires an infection while hospitalized. The variables include: infection risk (in percent), average length of patient stay (days),  average patient age(years), measure of how many x-rays are given in the hospital, and others.

(Data source: Applied Regression Models, (4th edition), Kutner, Neter, and Nachtsheim; it is in the file "hospital_infct.txt" available in the resource section of the course at Piazza.)

### Task 1
* Read-in the data and look at the several first lines of the data using head() command
* Run the linear regression of infection risk on all available predictors and print the summary.
* What are your conclusions? Which predictors are useful? How well does the regression fit the data?
 
```{r}
#riskdata = read(...)

#head(...)

#model.full = lm(...)
#summary(...)
```
You answer is here. 

### Task 2

Now run the regression on a smaller set of predictors, namely, Stay, Culture and Xray, and use the results to predict the infection risk for a hospital for which the average length of stay is 10 days, Culture = 14 and Xray = 82.

What is the prediction interval?

```{r}
#model.reduced = lm(...)
#summary(...)

#predict(...)
```
Your answer is herre

### Task 3
Compare the full and reduced model by calculating the F-statistic and the relevant p-value. 
What is your conclusion?

```{r}


```
Your answer is here.

### Task 4

This is an exercise about the matrix formulation of the regression model. 
Continue to use the model with variables Stay, Culture, Xray

* Construct y and X. (This can be done either directly or by using options 
y = T, x = T in the lm function).
* Calculate the coefficients of the regression by matrix operations 
* Calculate the prediction for Stay =10,  Culture = 14 and Xray = 82.

```{r}
#Extract y and X
#model = lm(..., x = T, y = T)
#X = ...
#y = ...

#calculate beta 
#beta = solve(...)
#beta

#calculate the prediction as $x.new^t \beta$ 
#x.new = c(....)
#y.new = ....
#y.new
```
 

