---
title: "Math 448 HW2"
author: "Vladislav Kargin"
date: '2022-09-18'
output: html_document 
---
 <!-- change pdf_document to html_document if you don't have TeX installed on your computer  -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HW2 

- You may discuss homework problems with other students, but **you have to prepare the written assignments yourself.**
- You can use the suggested outline for the R code or modify it in any way you like, or, if you prefer, write the code in an entirely different form that still accomplishes the required task.

### Problem 1

Let $X_1$, $X_2$, $X_3$, ... , $X_n$ be i.i.d exponential random variables with mean $\theta$.

1. One possible estimator for $\theta$ is $\overline X$. Recall that the variance of $X_i$ is $\theta^2$, so we can estimate $Var(X_i)$ as $(\overline X) ^2$. By using these estimators for the mean and variance, write down the asymptotic  (i.e., large sample) confidence interval for $\theta$ with the confidence level $1 - \alpha$. (Note that you should use $\overline X ^2$ as the variance estimator, not $S^2$.)



 Solution:
\[
\Big( \ldots \Big), 
\] 

2. Set $n = 4$, $\theta = 2$, $\alpha = 0.1$. Generate a vector $x$ that contains the random sample $X_1$, ... , $X_n$ and calculate the bounds of the asymptotic confidence interval which you wrote in 1. (Be careful with the generation of exponential r.v.s in R, since R functions for these random variables use parameter "rate" = $1/\theta$.) Does this interval contain $\theta$? 

Solution:
```{r}
set.seed(123)
# n = 
# theta = 
# alpha = 
# x = rexp(...)
#now calculate the lower and upper bounds of the confidence interval (lb, ub):
# lb = ... 
# lb
# ub =  ...
# ub
```
Comment: $\ldots$

3. Repeat the calculation that you programmed for 2. for $N = 10,000$ times and record whether the CI contained $\theta$ in a vector of length $N$. What is the proportion of the cases when $\theta$ lies inside the interval? Comment on your results: What can you say about the performance of the asymptotic CI? Why does this happen?

Solution:
```{r}
N = 10000
y = rep(0, N) #pre-fill y with zeros.
#for (i in 1:N){
#  ... #generate a sample and calculate the lower and upper bounds of the CI.
#  if (... && ...){ #if theta is between bounds, set y[i] to 1.
#    y[i] = ...
#  }
#}
# finally, calculate the mean of y
#  ...
```
Comment:  $\ldots$


4. Change $n$ to $n = 40$ and repeat 3. Comment on your results. 

Solution:
```{r}
#n = ...
#y = rep(...)
#for (...){
# ....   
#  if (...){
#    y[i] = ...
#  }
#}
# calculate the mean of y
# ...
```
Comment:  $\ldots$



